---
title: "R code for estimating the toxicokinetic parameters (*k*~u~ and *k*~e~)"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(deSolve) # for using the function "ode" to integrate
library(FME) # for using the function "modFit" to fit 
library(readxl) # for importing data from Excel files
library(ggplot2) # for plotting
library(dplyr) # for manipulating data frames
```

## 1. Cd isotope concentrations in exposure solution  

The data on Cd isotope concentrations in exposure solution were imported from the Excel file provided as another Supporting Information document. The *.xlsx file name in the code below should be revised to be the same as the that of the Excel file.
```{r message=FALSE, warning=FALSE}
d_water <-
  read_excel("SI_double isotope data.xlsx", sheet = 2) %>%
  filter(species == "X. atratus") 

```

## 2. Cd isotope concentrations in mussels  

Similarly, the data on the Cd isotope concentrations in the mussels during the double isotope exposure were imported from the same Excel file mentioned above but from a different sheet.
```{r message=FALSE, warning=FALSE}
d_organism <- read_excel("SI_double isotope data.xlsx", sheet = 1) %>%
  filter(species == "X. atratus")

```

## 3. Time points  

In the doulbe Cd isotope exposure of *X. atratus*, the T1 and T2 were 48 h and 336 h, respectively.  
```{r}
T1 <- 48 # T1 = 10 h
T2 <- 336 # T2 = 178
```

The total exposure length (i.e., T2 = 336) was divided into intervals of 1/48 d (i.e., 0.5 h). The time sequence was used below to interpolate the exposure concentration and calculate the bioaccumulation of Cd isotopes.
```{r}
times <- seq(from = 0,
             to = T2 / 24,
             by = 1 / 48)
```


## 4. Interpolation of Cd isotope concentrations in solutions  

Cd isotope concentrations in the exposure solutions during the whole experiment period was interpolated from the measured concentrations using the function `approxfun`.
```{r message=FALSE, warning=FALSE}
Cw_Cd113 <- approxfun(x = d_water$day,
                      y = d_water$Cw_Cd113_ppb,
                      rule = 2)
```


The interpolated concentrations of ^113^Cd are shown below:  

```{r message=FALSE, warning=FALSE}

data.frame(Day = times, Cd113 = Cw_Cd113(times)) %>%
  ggplot(aes(Day, Cd113)) +
  theme_classic() +
  geom_line(color = "red2") +
  labs(x = "Days",
       y = ~ "Concentration of " * "" ^ "113" * "Cd (" * mu * g ~ "L" ^
         "-1" * ")")
```

The interpolated concentrations of ^114^Cd are shown below:  

```{r message=FALSE, warning=FALSE}
Cw_Cd114 <- approxfun(x = d_water$day,
                      y = d_water$Cw_Cd114_ppb,
                      rule = 2)

data.frame(Day = times, Cd114 = Cw_Cd114(times)) %>%
  ggplot(aes(Day, Cd114)) +
  theme_classic() +
  geom_line(color = "red2") +
  scale_y_continuous(limits = c(0, NA)) +
  labs(x = "Days",
       y = ~ "Concentration of " * "" ^ "114" * "Cd (" * mu * g ~ "L" ^
         "-1" * ")")
```


## 5. Toxicokinetic (TK) model  

The toxicokinetic model is defined below. The variation of the mass of a Cd isotope in a mussel is the balance between uptake ($k_\text{in}C_\text{w}$) and elimination ($k_\text{e}M_\text{int}$) of the isotope. See equations 1-2 in the main text for detailed descriptions of the model.  

```{r message=FALSE, warning=FALSE}
TK <- function (t, y, parameters) {
  kin <- parameters[1]
  ke <- parameters[2]
  Mint_Cd113 <- y[1]
  Mint_Cd114 <- y[2]
  dMint_Cd113 <- kin * Cw_Cd113(t) - ke * Mint_Cd113
  dMint_Cd114 <- kin * Cw_Cd114(t) - ke * Mint_Cd114
  
  list(c(dMint_Cd113, dMint_Cd114))
}
```

## 6. Estimation of the initial parameter values  

The initial value of *k*~e~, over which the parameter value was optimized,  was estimated based on the decrease of ^113^Cd in mussels during the period between T1 and T2.  See Eq. 7 for more detailed description.

An overall rate of Cd elimination from the mussels can be estimated based on the decrease of 113Cd in mussels during the period between T1 and T2 (see the figure below):

$$ k_\text{e} = -\frac{\text{∆ln⁡(median of new }{}^{113} \text{Cd)}}{∆t}$$

where the numerator is the decrease of the median natural-log transformed ^113^Cd content, and the denominator *Δt* is the duration of the second stage (i.e., T2 – T1). 
    The overall *k*~e~ was used as the initial value to estimate *k*~e~ of individual mussels.     



```{r message=F}
ggplot(d_organism, aes(day, Cd113_new_ng)) +
  theme_classic() +
  geom_jitter(width = 0.1, height = 0, shape = 1) +
  stat_summary(geom = "point", fun = "median", color = "red") +
  stat_summary(geom = "line", fun = "median", color = "red") +
  labs(
    x = "Days",
     y = ~ "Content of " * "" ^ "113" * "Cd in soft tissue (ng)"
  )

```


```{r message=F}
d_organism_median <- d_organism %>%
  group_by(day) %>%
  summarise(median_Cd113 = median(Cd113_new_ng))

ke_initial <- with(d_organism_median,
                   (log(median_Cd113[2]) - log(median_Cd113[3])) / (day[3] -
                                                                  day[2]))
ke_initial 

```




## 7. Objective function for fitting toxicokinetic parameters  

The objective function returns a vector of the weighted residuals, which are the difference between the modeled amount of Cd isotopes and the measured amount of Cd isotopes. The reciprocal of the measured amount of Cd isotopes are used as the weight.  

```{r message=FALSE, warning=FALSE}
fn_resid <- function(parameters_to_fit) {
  kin <- parameters_to_fit[1]
  ke <- parameters_to_fit[2]
  out.Mint <- ode(
    func = TK,
    times = times,
    y = c(Mint_Cd113 = 0, Mint_Cd114 = 0),
    parms = parameters_to_fit
  )
  n <- length(times)
# residuals = (modeled – measured)/ measured
  resid.Mint <- c((out.Mint[n, 2] / organism_i$Cd113_new_ng - 1),
                  (out.Mint[n, 3] / organism_i$Cd114_new_ng - 1))
    return(resid.Mint)
}


```

## 8. Fitting *k*~in~ and *k*~e~ 

Cd isotope contents measured at time T2 are used for estimating *k*~in~ and *k*~e~.
```{r message=FALSE}
# selecting data of time T2
organism_T2 <- subset(d_organism, hour == T2) 
```


```{r message=FALSE}
# the number of organisms analyzed at T2
n_organism <- dim(organism_T2)[1] 

# empty vectors for storing kin and ke values
kin <- numeric(n_organism)
ke  <- numeric(n_organism)

# create a folder to store figures for checking the goodness of fitting.
fit_SN <- gsub("[: -]", "" , Sys.time())
path <- paste0("fit_check_", fit_SN)
dir.create(path)
```

Fitting the values of *k*~in~ and *k*~e~ for each individual. This step may take 5 to 10 min to finish.
```{r message=FALSE}

for (i in 1:n_organism) {
  # select data for the ith individual
  organism_i <- organism_T2[i, ]
  # intial estimation of kin: content of 114Cd/exposure time/exposure concentration  
  kin_initial <-
    organism_i$Cd114_new_ng / (T2 / 24) / mean(d_water$Cw_Cd114_ppb)

print(system.time(
  # using the optimization function `modFit` to estimate the values of kin and ke
  fit.TK <- modFit(
    # objective function defined above
    f = fn_resid,
    # initial parameter values estimated above
    p = c(kin_initial, ke_initial),
    # the quasi-Newton method
    method = "BFGS",
    # the lower and upper bound of kin and ke
    lower = c(kin_initial * 0.1, ke_initial * 0.1),
    upper = c(kin_initial * 10, ke_initial * 10)
  )
))

# extract parameter values and store them into the corresponding vectors
kin[i] <- fit.TK$par[1]
ke[i]  <- fit.TK$par[2]

# extract parameter values for the integration below
kin_i <- fit.TK$par[1]
ke_i  <- fit.TK$par[2]

# integrate the TK model to calculate content of Cd isotopes at the end of exposure
fit_i <- ode(
  func = TK,
  times = times,
  y = c(Cint_113 = 0, Cint_114 = 0),
  parms = c(kin_i, ke_i)
)
d_fit_i <- as.data.frame(fit_i)
# plot the modeled content against the measured content to check the goodness of fitting  
ggplot(d_fit_i, aes(x = time, Cint_113)) +
  geom_line() +
  geom_line(aes(x = time, y = Cint_114), color = "red2") +
  geom_point(aes(x = organism_i$day, organism_i$Cd113_new_ng)) +
  geom_point(aes(x = organism_i$day, organism_i$Cd114_new_ng), color =
               "red2")

ggsave(paste("fit", i, ".png"), path = path)

}


```


## 9. Save *k*~u~，*k*~e~   

The values of *k*~u~，*k*~e~ are saved as a *.csv file.  
```{r message=FALSE, warning=FALSE}
organism_T2$kin <- kin
# ku is calculated from kin: ku = kin / weight of organism
organism_T2$ku <- kin / 1000 / organism_T2$dw_g
organism_T2$ke  <- ke
# save the values as a csv file 
write.csv(organism_T2, file=paste0("fitted parameters_", fit_SN, ".csv"))

```

## 10. Goodness of fit  

The goodness of fit is visualized by plotting the modeled content vs. the measured content of ^113^Cd and ^114^Cd in each organism.  

```{r message=FALSE, warning=FALSE}
# empty vector to store modeled content of Cd isotopes  
Cd113_fit <- numeric(n_organism)
Cd114_fit <- numeric(n_organism)

# Cd isotope content calculated by integrating the TK model  
for (i in 1:n_organism){  
  out.fit <- ode(func = TK, 
                 times = times,
                 y = c(Cint_Cd113 = 0, Cint_Cd114 = 0), 
                 parms = c(kin = kin[i], ke = ke[i]))
  N_step <- length(times)
  Cd113_fit[i] <- out.fit[N_step, 2]
  Cd114_fit[i] <- out.fit[N_step, 3]
}

organism_T2$Cd113_fit <- Cd113_fit
organism_T2$Cd114_fit <- Cd114_fit

# save the modeled Cd content as a csv file  
write.csv(organism_T2, 
          file=paste0("fit_Cd_T2_",fit_SN,".csv"))

# plot the modeled contet vs. measured content
ggplot(organism_T2, aes(Cd113_new_ng, Cd113_fit))+
  theme_classic()+
  geom_point(aes(color = "Cd-113"))+
  geom_point(aes(Cd114_new_ng, Cd114_fit, color = "Cd-114"))+
  geom_abline(slope = 1, intercept = 0)+
  scale_x_log10()+
  scale_y_log10()+
  scale_color_manual(values = c("black", "red2")) +
  labs(x = "Measured newly accumulated Cd (ng)",
       y = "Modeled newly accumulated Cd (ng)",
       color = NULL)

```